Single-fluorophore orientation determination with multiview polarized illumination: modeling and microscope design

### 2017-09-05 paper/paper/paper.tex
Round of small edits from all authors.

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
%% Comment before submission
\bibliography{paper}
\bibliographystyle{osajnl}

@@

@@

\textbf{Polarized illumination or detection} techniques vary the polarization of
light in the illumination or detection paths and exploit the anisotropic
absorption and emission patterns of single molecules to estimate their
orientation. By illuminating and detecting from the entire sample, not just the
focal volume, these techniques can estimate the orientation of many molecules in
parallel without scanning. These techniques are easy to implement---changing the
illumination or detection polarization is simple with a polarization splitter
\cite{mehta2016} or universal compensator/polarizer \cite{shribak2003}---and the
reconstruction methods are straightforward \cite{fourkas2001, mehta2016,
  backer2016}. Also, these techniques keep the signal from a single molecule
localized on the detector, so they have a signal-to-noise ratio advantage over
defocusing and phase masking techniques. The main drawback is that these
techniques require several intensity measurements under different polarization
orientations to estimate the orientation of a single molecule. Despite the
requirement for multiple measurements, polarized illumination or detection
methods are good choices for live cell imaging because of their combination of
speed and ease of implementation.

@@

In this work we explore the use of polarized illumination in multiview
microscopes for determining the orientation of single molecules. Existing
multiview microscopes can easily be outfitted with fast-switching polarizers
that do not degrade image quality, so polarized illumination is a natural way to
augment multiview microscopes for measuring the orientation of single
molecules. Furthermore, multiview microscopes can achieve nearly isotropic
resolution while delivering selective illumination. We find that multiview
microscopes also provide a small and uniform orientation uncertainty for single
molecules in all orientations. We also find that single-view microscopes have
symmetries that can be broken by adding an second view. This reduction of
degeneracy allows multiview microscopes to uniquely determine the orientation of
single molecules.

@@

Note that $A$ and $B$ in equation \ref{eq:coefficients} are a factor of
$\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
Fourkas' expressions were incorrectly normalized (the limit of
$\eta_{\text{det}}$ as $\alpha\rightarrow \pi/2$ should be $\frac{1}{2}$), and
the extra factor of $\frac{3}{2}$ corrects the error. This correction does not
affect Fourkas' orientation reconstruction, but it does mean that Fourkas'
algorithm underpredicts the total emitted intensity by a factor of
$\frac{3}{2}$.

Also note that the detection efficiency can take its maximum range of values
when $B$ takes its maximum value, i.e. when
$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When
$\alpha = 90$ the detection efficiency is $\frac{1}{2}$ for all fluorophore
orientations and when $\alpha = 0$ the detection efficiency is zero for all
fluorophore orientations, so we expect there to be an intermediate value of
$\alpha$ where the detection efficiency takes its maximum range.

\subsection{Oblique Illumination and Detection}\label{oblique}
In sections \ref{excitation} and \ref{detection} we assumed that both the
illumination and detection objectives had $\mh{z}$ aligned optical axes, i.e.,
we used epi- or transmitted light illumination. To extend the forward model to
oblique optical axes which allows for non-coincident illumination and detection
objectives, we express the dipole orientation in a rotated coordinate frame
using the following expressions
\begin{align}
    \Theta' &= \arccos\left(\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta\right)\label{eq:thetap}\\
  \Phi' &=
          \begin{cases}
            \arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &0 \leq \Phi < \pi  \\
            -\arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &-\pi \leq \Phi < 0\\
          \end{cases}\label{eq:phip}
\end{align}
where $\psi$ is the angle of a right handed rotation about the $\mh{y}$ axis
that maps the $\mh{z}$ axis onto the new optical axis.

\subsection{Intensity Measurements}\label{forward}
The detected intensity is proportional to the product of the sample exposure
$I_{\text{exp}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I \propto I_{\text{exp}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
\end{align}
where the proportionality depends on the exposure time, the molecular quantum
efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
forward model for a single intensity measurement. Figure
\ref{fig:single-frame} shows the efficiencies in three representative
geometries. Note that in all cases there are multiple orientations that give the
same intensity measurement. This means that a single intensity measurement does
not give us enough information to reconstruct the orientation of a single
molecule. Next, we will consider combining several intensity measurements to
create complete polarized illumination microscopes.

@@

For each microscope we calculate $\sigma_{\Omega}$ at 10,000 approximately
equally spaced points on the unit sphere. A desirable microscope design will
have a small solid-angle uncertainty that is uniform for all molecule
orientations. The most straightforward way to find the location and scale of the
solid-angle uncertainty is to use the mean and variance, respectively. However,
the solid-angle uncertainty can take extremely large values, e.g., when a
fluorescent dipole is aligned with the optical axis of a single excitation
objective. Such large values can change the mean and variance dramatically
depending on the sample points, so we use the median and median absolute
deviation (MAD)---the median of the data's absolute difference from the
median---as robust alternatives to the mean and variance.

@@

The solid-angle uncertainty and the uncertainty ellipse are good local measures
of orientation uncertainty, but they are insensitive to degeneracies with
distant dipole orientations. For example, all of the intensity measurements in
this paper are identical under inversion of the dipole orientation
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in equations \ref{eq:scalarabs},
\ref{eq:scalardet}, and \ref{eq:forward_frame}), but the solid-angle uncertainty
and the uncertainty ellipse give no indication of this degeneracy. For this
reason we also use the number of degeneracies as a measure of a micrsocope's
ability to estimate the orientation of a dipole from intensity measurements. A
small number of degeneracies is important for a microscope to find the
orientation of a molecule---a twofold degenerate microscope can find the
orientation of a molecule down to the hemisphere while a fourfold symmetric
microscope can only find the orientation down to the quadrant.

@@

The single-view microscope in Figure \ref{fig:single-arm} is fourfold degenerate
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
\pi, \Phi+\pi)$ in equations \ref{eq:scalarabs}, \ref{eq:scalardet}, and
\ref{eq:forward_frame}). In addition to the usual inversion degeneracy,
single-view microscopes suffer from a degeneracy when a dipole is reflected
through the transverse plane. Evidence of this degeneracy is found in Figures
\ref{fig:single-arm}b) and \ref{fig:single-arm}c) which show the ambiguity of
whether the molecule is oriented above or below the transverse plane. Near the
transverse plane this ambiguity is evident in the high solid-angle uncertainty,
but far from the transverse plane evidence of the ambiguity is lost. Adding a
second view and taking eight intensity measurements instead of four removes this
degeneracy (see equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view
microscopes we consider in this paper are only twofold degenerate. This means
that while dual-view microscopes can find the orientation of a molecule up to
the hemisphere, while single-view microscopes can only find the orientation of a
molecule up to the quadrant.

@@

Table 1 shows a summary of our evaluation metrics for microscopes in each
class. The single-view microscope with a small illumination NA and high
detection NA has the lowest solid-angle uncertainty median but the largest max
and fourfold symmetry. The dual-view oblique symmetric with an
intermediate NA performs reasonably well on all three metrics, while the
dual-view orthogonal symmetric light-sheet microscopes with large NA perform
very well on all three metrics.

@@

\section*{Funding}
Support for this work was provided by the Intramural Research Programs of the
National Institute of Biomedical Imaging and Bioengineering.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
%% Comment before submission
% \bibliography{paper}
% \bibliographystyle{osajnl}

@@

\begin{thebibliography}{10}
\newcommand{\enquote}[1]{``#1''}

\bibitem{peterman2001}
E.~J. Peterman, H.~Sosa, L.~S. Goldstein, and W.~E. Moerner, \enquote{Polarized
  fluorescence microscopy of individual and many kinesin motors bound to
  axonemal microtubules,} Biophys. J. \textbf{81}, 2851--2863 (2001).

\bibitem{forkey2003}
J.~N. Forkey, M.~E. Quinlan, M.~A. Shaw, J.~E.~T. Corrie, and Y.~E. Goldman,
  \enquote{Three-dimensional structural dynamics of myosin {V} by
  single-molecule fluorescence polarization,} \nat \textbf{422}, 399--404
  (2003).

\bibitem{backer2016}
A.~S. Backer, M.~Lee, and W.~E. Moerner, \enquote{Enhanced {DNA} imaging using
  super-resolution microscopy and simultaneous single-molecule orientation
  measurements,} in \enquote{Conference on Lasers and Electro-Optics,}
  (Optical Society of America, 2016), p. JTh4B.4.

\bibitem{mehta2016}
S.~B. Mehta, M.~McQuilken, P.~J. La~Rivi\`ere, P.~Occhipinti, A.~Verma,
  R.~Oldenbourg, A.~S. Gladfelter, and T.~Tani, \enquote{Dissection of
  molecular assembly dynamics by tracking orientation and position of single
  molecules in live cells,} Proc. Natl. Acad. Sci. U.S.A. \textbf{113},
  E6352--E6361 (2016).

\bibitem{demay2011}
B.~DeMay, N.~Noda, A.~Gladfelter, and R.~Oldenbourg, \enquote{Rapid and
  quantitative imaging of excitation polarized fluorescence reveals ordered
  septin dynamics in live yeast,} Biophys. J. \textbf{101}, 985--994 (2011).

\bibitem{mcquilken2017}
M.~McQuilken, M.~S. Jentzsch, A.~Verma, S.~B. Mehta, R.~Oldenbourg, and A.~S.
  Gladfelter, \enquote{Analysis of septin reorganization at cytokinesis using
  polarized fluorescence microscopy,} Front. Cell Dev. Biol. \textbf{5}, 42
  (2017).

\bibitem{anantharam2010}
A.~Anantharam, B.~Onoa, R.~H. Edwards, R.~W. Holz, and D.~Axelrod,
  \enquote{Localized topological changes of the plasma membrane upon exocytosis
  visualized by polarized {TIRFM},} J. Cell Biol. \textbf{188}, 415--428
  (2010).

\bibitem{backlund2014}
M.~P. Backlund, M.~D. Lew, A.~S. Backer, S.~J. Sahl, and W.~E. Moerner,
  \enquote{The role of molecular dipole orientation in single-molecule
  fluorescence microscopy and implications for super-resolution imaging,}
  ChemPhysChem \textbf{15}, 587--599 (2014).

\bibitem{lieb2004}
M.~A. Lieb, J.~M. Zavislan, and L.~Novotny, \enquote{Single-molecule
  orientations determined by direct emission pattern imaging,} J. Opt. Soc. Am.
  B \textbf{21}, 1210--1215 (2004).

\bibitem{backer2014}
A.~S. Backer and W.~E. Moerner, \enquote{Extending single-molecule microscopy
  using optical {F}ourier processing,} J. Phys. Chem. B \textbf{118},
  8313--8329 (2014).

\bibitem{agrawal2012}
A.~Agrawal, S.~Quirin, G.~Grover, and R.~Piestun, \enquote{Limits of 3{D}
  dipole localization and orientation estimation for single-molecule imaging:
  towards {G}reen's tensor engineering,} Opt. Express \textbf{20}, 26667--26680
  (2012).

\bibitem{toprak2006}
E.~Toprak, J.~Enderlein, S.~Syed, S.~A. McKinney, R.~G. Petschek, T.~Ha, Y.~E.
  Goldman, and P.~R. Selvin, \enquote{Defocused orientation and position
  imaging ({DOPI}) of myosin {V},} Proc. Natl. Acad. Sci. U.S.A. \textbf{103}, 6495--6499 (2006).

\bibitem{debarre2004}
A.~D{\'e}barre, R.~Jaffiol, C.~Julien, D.~Nutarelli, A.~Richard,
  P.~Tch{\'e}nio, F.~Chaput, and J.-P. Boilot, \enquote{Quantitative
  determination of the 3{D} dipole orientation of single molecules,} Eur. Phys.
  J. D. \textbf{28}, 67--77 (2004).

\bibitem{shribak2003}
M.~Shribak and R.~Oldenbourg, \enquote{Techniques for fast and sensitive
  measurements of two-dimensional birefringence distributions,} Appl. Opt.
  \textbf{42}, 3009--3017 (2003).

\bibitem{fourkas2001}
J.~T. Fourkas, \enquote{Rapid determination of the three-dimensional
  orientation of single molecules,} Opt. Lett. \textbf{26}, 211--213 (2001).

\bibitem{tyo2006}
J.~S. Tyo, D.~L. Goldstein, D.~B. Chenault, and J.~A. Shaw, \enquote{Review of
  passive imaging polarimetry for remote sensing applications,} Appl. Opt.
  \textbf{45}, 5453--5469 (2006).

\bibitem{lu2008}
C.-Y. Lu and D.~A.~V. Bout, \enquote{Analysis of orientational dynamics of
  single fluorophore trajectories from three-angle polarization experiments,}
  J. Chem. Phys. \textbf{128}, 244501 (2008).

\bibitem{wu2013}
Y.~Wu, P.~Wawrzusin, J.~Senseney, R.~S. Fischer, R.~Christensen, A.~Santella,
  A.~G. York, P.~W. Winter, C.~M. Waterman, Z.~Bao, D.~A. Colon-Ramos,
  M.~McAuliffe, and H.~Shroff, \enquote{Spatially isotropic four-dimensional
  imaging with dual-view plane illumination microscopy,} Nat. Biotechnol.
  \textbf{31}, 1032--1038 (2013).

\bibitem{keller2015}
R.~K. Chhetri, F.~Amat, Y.~Wan, B.~Hockendorf, W.~C. Lemon, and P.~J. Keller,
  \enquote{Whole-animal functional and developmental imaging with isotropic
  spatial resolution,} Nat. Methods \textbf{12}, 1171--1178 (2015).

\bibitem{wu2016}
Y.~Wu, P.~Chandris, P.~W. Winter, E.~Y. Kim, V.~Jaumouill{\'e}, A.~Kumar,
  M.~Guo, J.~M. Leung, C.~Smith, I.~Rey-Suarez, H.~Liu, C.~M. Waterman, K.~S.
  Ramamurthi, P.~J. La~Rivi\`ere, and H.~Shroff, \enquote{Simultaneous
  multiview capture and fusion improves spatial resolution in wide-field and
  light-sheet microscopy,} Optica \textbf{3}, 897--910 (2016).

\bibitem{wu2017}
Y.~Wu, A.~Kumar, C.~Smith, E.~Ardiel, P.~Chandris, R.~Christensen,
  I.~Rey-Suarez, M.~Guo, H.~Vishwasrao, J.~Chen, J.~Tang, A.~Upadhyaya,
  P.~La~Rivi\`ere, and H.~Shroff, \enquote{Reflective imaging improves
  resolution, speed, and collection efficiency in light sheet microscopy,}
  bioRxiv  (2017).

\bibitem{kay1993}
S.~M. Kay, \emph{Fundamentals of statistical signal processing} (Prentice Hall
  PTR, 1993).

\bibitem{anderson1958}
T.~W. Anderson, \emph{An introduction to multivariate statistical analysis}
  (Wiley New York, 1958).

\bibitem{coe2009}
D.~Coe, \enquote{Fisher matrices and confidence ellipses: a quick-start guide
  and software,} arXiv preprint arXiv:0906.4123  (2009).

\bibitem{walt2011}
S.~van~der Walt, S.~C. Colbert, and G.~Varoquaux, \enquote{The {N}um{P}y array:
  a structure for efficient numerical computation,} CoRR \textbf{abs/1102.1523}
  (2011).

\bibitem{meurer2017}
A.~Meurer, C.~P. Smith, M.~Paprocki, O.~\v{C}ert\'{i}k, S.~B. Kirpichev,
  M.~Rocklin, A.~Kumar, S.~Ivanov, J.~K. Moore, S.~Singh, T.~Rathnayake,
  S.~Vig, B.~E. Granger, R.~P. Muller, F.~Bonazzi, H.~Gupta, S.~Vats,
  F.~Johansson, F.~Pedregosa, M.~J. Curry, A.~R. Terrel, v.~Rou\v{c}ka,
  A.~Saboo, I.~Fernando, S.~Kulal, R.~Cimrman, and A.~Scopatz,
  \enquote{Sym{P}y: symbolic computing in {P}ython,} PeerJ Computer Science
  \textbf{3}, e103 (2017).

\bibitem{hunter2007}
J.~D. Hunter, \enquote{Matplotlib: A {2D} graphics environment,} Computing In
  Science \& Engineering \textbf{9}, 90--95 (2007).

\bibitem{vispy}
L.~Campagnola, A.~Klein, E.~Larson, A.~Griffiths, C.~Rossant, N.~P. Rougier,
  I.~Zaid, M.~S. Suraj, A.~Taylor, sylm21, M.~F. Kaptan, A.~J. Champandard,
  karlcz, nippoo, Siddharth, X.~Gao, C.~GESTES, bdurin, J.~D. Reaver,
  R.~McMullen, FiachAntaw, T.~Mansencal, M.~Dartiailh, StephenWest, J.~Kieffer,
  G.~Baty, R.~Peters, macduff111, E.~S. de~Andrade, and D.~Rufat,
  \enquote{vispy: Version 0.4.0,}  (2015).

\bibitem{bowman2008}
J.~C. Bowman and A.~Hammerlindl, \enquote{Asymptote: A vector graphics
  language,} TUGBOAT \textbf{29}, 288--294 (2008).

\bibitem{nov2006}
L.~Novotny and B.~Hecht, \emph{Principles of Nano-Optics} (Cambridge University
  Press, 2006).

\bibitem{levoy2006}
M.~Levoy, R.~Ng, A.~Adams, M.~Footer, and M.~Horowitz, \enquote{Light field
  microscopy,} in \enquote{ACM SIGGRAPH 2006 Papers,}  (ACM, New York, NY, USA,
  2006), SIGGRAPH '06, pp. 924--934.

\end{thebibliography}

@@

\textbf{Polarized illumination or detection} techniques vary the polarization of
light in the illumination or detection paths and exploit the anisotropic
absorption and emission patterns of single molecules to estimate their
orientation. By illuminating and detecting from the focal plane, not just the
focal point, these techniques can estimate the orientation of many molecules in
parallel without scanning. These techniques are easy to implement---changing the
illumination or detection polarization is simple with a polarization splitter
\cite{mehta2016} or universal compensator/polarizer \cite{shribak2003}---and the
reconstruction methods are straightforward \cite{fourkas2001, mehta2016,
  backer2016}. Also, these techniques keep the signal from a single molecule
localized on the detector, so they have a signal-to-noise ratio advantage over
defocusing and phase masking techniques. The main drawback is that these
techniques require several intensity measurements under different polarization
orientations to estimate the orientation of a single molecule. Despite the
requirement for multiple measurements, polarized illumination or detection
methods are good choices for live cell imaging because of their combination of
speed and ease of implementation.

@@

In this work we explore the use of polarized illumination in multiview
microscopes for determining the orientation of single molecules. Existing
multiview microscopes can easily be outfitted with fast-switching polarizers
that do not degrade image quality, so polarized illumination is a natural way to
augment multiview microscopes for measuring the orientation of single
molecules. Furthermore, multiview microscopes can achieve nearly isotropic
resolution while delivering selective illumination. We find that multiview
microscopes also provide a small and uniform orientation uncertainty for single
molecules in all orientations. We also find that single-view microscopes have
symmetries that can be broken by adding a second view. This reduction of
degeneracy allows multiview microscopes to uniquely determine the orientation of
single molecules.

@@

Note that $A$ and $B$ in equation \ref{eq:coefficients} are a factor of
$\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
Fourkas' expressions were incorrectly normalized (the limit of
$\eta_{\text{det}}$ as $\alpha\rightarrow \frac{\pi}{2}$ should be $\frac{1}{2}$), and
the extra factor of $\frac{3}{2}$ corrects the error. This correction does not
affect Fourkas' orientation reconstruction, but it does mean that Fourkas'
algorithm underpredicts the total emitted intensity by a factor of
$\frac{3}{2}$.

Also note that the detection efficiency can take its maximum range of values
when $B$ takes its maximum value, i.e. when
$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When
$\alpha = \frac{\pi}{2}$ the detection efficiency is $\frac{1}{2}$ for all
fluorophore orientations and when $\alpha = 0$ the detection efficiency is zero
for all fluorophore orientations, so we expect there to be an intermediate value
of $\alpha$ where the detection efficiency takes its maximum range. This means
that the maximum modulation of detected intensities will occur when we choose a
detection objective with $\alpha \approx 54.7^{\circ}$.

\subsection{Oblique Illumination and Detection}\label{oblique}
In sections \ref{excitation} and \ref{detection} we assumed that both the
illumination and detection objectives had $\mh{z}$ aligned optical axes, i.e.,
we assumed traditional epi- or transmitted light illumination. To extend the
forward model to oblique optical axes which allows for non-coincident
illumination and detection objectives, we express the dipole orientation in a
rotated coordinate frame using the following expressions
\begin{align}
    \Theta' &= \arccos\left(\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta\right)\label{eq:thetap}\\
  \Phi' &=
          \begin{cases}
            \arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &0 \leq \Phi < \pi  \\
            -\arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &-\pi \leq \Phi < 0\\
          \end{cases}\label{eq:phip}
\end{align}
where $\psi$ is the angle of a right handed rotation about the $\mh{y}$ axis
that maps the $\mh{z}$ axis onto the new optical axis.

\subsection{Intensity Measurements}\label{forward}
The detected intensity is proportional to the product of the sample exposure
$I_{\text{exp}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I \propto I_{\text{exp}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
\end{align}
where the proportionality depends on the exposure time, the molecular quantum
efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
forward model for a single intensity measurement. Figure \ref{fig:single-frame}
shows the efficiencies in three representative geometries. Note that in all
cases there are multiple orientations that give the same intensity
measurement. This means that a single intensity measurement does not give us
enough information to reconstruct the orientation of a single molecule. Next, we
will consider combining several intensity measurements to create polarized
illumination microscopes that can reconstruct the orientation of single
molecules.

@@

For each microscope we calculate $\sigma_{\Omega}$ at 10,000 approximately
equally spaced points on the unit sphere. A desirable microscope design will
have a small solid-angle uncertainty that is uniform for all molecular
orientations. The most straightforward way to find the location and scale of the
solid-angle uncertainty is to use the mean and variance, respectively. However,
the solid-angle uncertainty can take extremely large values, e.g., when a
fluorescent dipole is aligned with the optical axis of a single excitation
objective. Such large values can change the mean and variance dramatically
depending on the sample points, so we use the median and median absolute
deviation (MAD)---the median of the data's absolute difference from the
median---as robust alternatives to the mean and variance.

@@

The solid-angle uncertainty and the uncertainty ellipse are good local measures
of orientation uncertainty, but they are insensitive to degeneracies with
distant dipole orientations. For example, all of the intensity measurements in
this paper are identical under inversion of the dipole orientation
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in equations \ref{eq:scalarabs},
\ref{eq:scalardet}, and \ref{eq:forward_frame}), but the solid-angle uncertainty
and the uncertainty ellipse give no indication of this degeneracy. For this
reason we also use the number of degeneracies as a measure of a microscope's
ability to estimate the orientation of a dipole from intensity measurements. A
small number of degeneracies is important for a microscope to find the
orientation of a molecule---a twofold degenerate microscope can find the
orientation of a molecule down to the hemisphere while a fourfold degenerate
microscope can only find the orientation down to the quadrant.

@@

The single-view microscope in Figure \ref{fig:single-arm} is fourfold degenerate
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
\pi, \Phi+\pi)$ in equations \ref{eq:scalarabs}, \ref{eq:scalardet}, and
\ref{eq:forward_frame}). The first twofold degeneracy is due to the usual
inversion degeneracy---the dipole is an axis rather than a direction. A second
twofold degeneracy is present in single-view designs---the microscope cannot
distinguish between dipoles reflected through the transverse plane. Evidence of
the second degeneracy is found in Figures \ref{fig:single-arm}b) and
\ref{fig:single-arm}c) which show the ambiguity of whether the molecule is
oriented above or below the transverse plane. Near the transverse plane this
ambiguity is evident in the high solid-angle uncertainty, but far from the
transverse plane evidence of the ambiguity is lost. Adding a second view and
taking eight intensity measurements instead of four removes this degeneracy (see
equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view microscopes we
consider in this paper are only twofold degenerate. This means that while
dual-view microscopes can find the orientation of a molecule up to the
hemisphere, single-view microscopes can only find the orientation of a molecule
up to the quadrant.

@@

Table 1 shows a summary of our evaluation metrics for microscopes in each
class. The single-view microscope with a small illumination NA and high
detection NA has the lowest solid-angle uncertainty median but the largest max
and fourfold degeneracy. The dual-view oblique symmetric microscope with an
intermediate NA performs reasonably well on all three metrics, while the
dual-view orthogonal symmetric light-sheet microscopes with large NA perform
very well on all three metrics.

@@

\section*{Funding}
TC was supported by a University of Chicago Biological Sciences Division
Graduate Fellowship. SM and RO were supported by NIH grant R01GM114274. HS was
supported by the Intramural Research Programs of the National Institute of
Biomedical Imaging and Bioengineering.
